---
title: "HW1_Aisha_Al-Khinji"
author: "Aisha Al-Khinji"
date: "01/27/2015"
output: html_document
---

This is my first homework 

```{r, echo=TRUE}
library(ggplot2)
```

Load the iris data

```{r, echo=TRUE}
data(iris)
```

using the variable directly

```{r, echo=TRUE}
qplot(iris$Sepal.Length, iris$Sepal.Width)
```
  
use the dataframe

```{r, echo=TRUE}
qplot(Sepal.Length, Sepal.Width, data=iris)
```

```{r, echo=TRUE}
head(iris)
```

Load the library

```{r, echo=TRUE}
library(reshape2)
```

reshape2-melt

add a columen to keep track of the flower


```{r, echo=TRUE}
iris$flower_id <- rownames(iris)
```

Defalult options

```{r, echo=TRUE}
iris_melted <- melt(iris)
```

Using Species, flower_id as id variables

```{r, echo=TRUE}
head(iris_melted)
```

split the variable to get the columns as we need

```{r, echo=TRUE}
split_variable <- strsplit(as.character(iris_melted$variable),split="\\.")
```

creat two new variables 

```{r, echo=TRUE}
iris_melted$flower_part <- sapply(split_variable,"[",1)
iris_melted$measurement_type <- sapply(split_variable,"[",2)
```

remove one of the variable that we don't need

```{r, echo=TRUE}
head(iris_melted)
```

reshape2-cast

use cast to frame the data 

```{r, echo=TRUE}
iris_cast <- dcast(iris_melted, formula=flower_id+Species+flower_part~measurement_type)
```

```{r, echo=TRUE}
head(iris_cast)
```

facet by species and Petal/Sepal

```{r, echo=TRUE}
qplot (x=Width, y=Length, data=iris_cast, geom=c("point", "smooth"), color=Species, method="lm", facets=flower_part~Species)
```

facet by flower part by using Species and aesthetic variable and using facet_wrap

```{r, echo=TRUE}
ggplot(data=iris_cast, aes(x=Width, y=Length))+geom_point()+facet_grid(Species~flower_part, scale="free")+geom_smooth(method="lm")+theme_bw(base_size=10)
```

